---
title: "coxph_present"
author: "Congyang Xie"
date: "`r Sys.Date()`"
output:
  beamer_presentation:
    colortheme: "dolphin"
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(survival)
library(survminer)
```

```{r include=FALSE}
data(cancer, package="survival")
```



```{r include=FALSE}
bladder1 <-
  bladder1 %>%
  mutate(recurr = if_else(status == 1, 1, 0),
         time = stop - start) %>%
  filter(start == 0)
```


## Fit coxph model

+ Fit a Cox proportional hazard model with initial tumor number and size as covariates.
```{r echo=FALSE}
bladder1.cox.2 <- coxph(Surv(time, recurr) ~ treatment + number + size, data = bladder1)
# summary(bladder1.cox.2)
bladder1.cox.2 %>%  
  broom::tidy() %>% 
  mutate(exp_est = exp(estimate)) %>% 
  relocate(term, estimate, exp_est) %>% 
  knitr::kable(digits = 3, caption = "Coefficients")
```

+ Treatment thiotepa and initial tumor number is statistically significant
+ The thiotepa treatment help decrease the hazard, thus improving the prognosis of bladder cancer
+ The initial tumor number increases the hazard, thus the prognosis worse, for subjects with more tumors at the beginning.
+ exp_est is the hazard ratio. Compared to the placebo, the thiotepa treatment reduce the hazard of 42.4%

## Fit coxph model
### Test statistics

```{r echo=FALSE}
stat.val <-
  bladder1.cox.2 %>%
  broom::glance() %>%
  select(starts_with("statistic")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "test_name",
    names_prefix = "statistic.",
    values_to = "statistic_value"
  ) %>%
  filter(test_name != "robust")

p.val <-
  bladder1.cox.2 %>%
  broom::glance() %>%
  select(starts_with("p.value")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "test_name",
    names_prefix = "p.value.",
    values_to = "p_value"
  ) %>%
  filter(test_name != "robust") %>%
  select(-test_name)

bind_cols(stat.val, p.val) %>% knitr::kable(digits = 3, caption = "Global statistical significance")
```
+ the output gives p-values for three alternative tests for overall significance of the model: The likelihood-ratio test, score log-rank statistics, and Wald test.
+ These three methods are asymptotically equivalent

## Check assumption

```{r echo=FALSE}
test.ph.2 <- cox.zph(bladder1.cox.2)
test.ph.2[["table"]] %>% knitr::kable(digits = 3)
```

+ From the output above, the test is not statistically significant for each of the covariates, and the global test is also not statistically significant. 
+ Therefore, we can assume the proportional hazards.


## Check assumption

```{r echo=FALSE, fig.height=4.5, fig.width=8}
ggcoxzph(test.ph.2)
```

+ Systematic departures from a horizontal line are indicative of non-proportional hazards, since proportional hazards assumes that estimates do not vary much over time.
+ From the graphical inspection, there is no pattern with time.

## check influential observation

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Testing influential observations
ggcoxdiagnostics(
  bladder1.cox.2,
  type = "dfbeta"
)
```

+ Plots the estimated changes in the regression coefficients upon deleting each observation in turn
+ The index plots above demonstrate that none of the observations are particularly influential on their own.

## Testing non linearity

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
# Testing non linearity
ggcoxfunctional(Surv(time, recurr) ~  number + log(number) + sqrt(number), data = bladder1)
```

+ The plot display graphs of continuous covariates against residuals of null cox proportional hazards model. 
+ It appears that, there's slight non-linearity for the initial tumor number here.

## Testing non linearity

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE}
ggcoxfunctional(Surv(time, recurr) ~  size + log(size) + sqrt(size), data = bladder1)
```

+ It appears that, there's non-linearity for the initial tumor size here.