---
title: "EDA"
author: "XIAO MA"
date: '2022-11-27'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggridges)
library(dplyr)
library(survival)
```

```{r}
theme_set(
  theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5),
      plot.caption = element_text(hjust = 0.0)
    )
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
data(cancer, package = "survival")
head(bladder1) %>% knitr::kable()
```

* id:	Patient id
* treatment:	Placebo, pyridoxine (vitamin B6), or thiotepa
* recur:	Number of recurrences
* start,stop:	The start and end time of each time interval
* status:	End of interval code, 0=censored, 1=recurrence, 2=death from bladder disease, 3=death other/unknown cause
* rtumor:	Number of tumors found at the time of a recurrence
* rsize:	Size of largest tumor at a recurrence
* enum:	Event number (observation number within patient)

```{r}
bladder = bladder1 %>% group_by(id) %>% 
  mutate(total_recs = n(),
         new_enum =-enum + total_recs + 1) %>% 
  filter(new_enum == 1)
```

- Summarize data (plots of tumor information, # patients assigned to each treatment, # of each status [censored, died, recurrence], etc.)

** Distribution of patients for Each Treatment and Recurrence 
```{r}
bladder %>% select(id,treatment,recur) %>% filter(recur > 0) %>%  group_by(treatment) %>% summarise(n = n()) 

bladder %>% select(id,treatment,recur) %>% filter(recur == 0) %>% unique() %>% group_by(treatment) %>% summarise(n = n())

p3 = data.frame(treatment = c(rep("placebo",2),rep("pyridoxine",2), rep("thiotepa",2)),
  recurrence = rep(c("Y","N"),3),n = c(29,19,15,17,18,20))

ggplot(data = p3, aes(x = treatment, y = n, fill=factor(recurrence))) + geom_bar(stat="identity",color = "black") + geom_text(aes(label = n),color = "black",hjust = 0.5,vjust = 3,size = 3,position = "stack") + labs(title = "Distribution of patients for Each Treatment and Recurrence ", x = "Treatment", y ="Number of Patients") + guides(fill = guide_legend(title = "Recurrence")) + scale_fill_manual(values = c('light blue','orchid'))

```
The number of patients in the three treatment groups are shown. The patients were assigned at random with equal probability but numbers in the three treatment groups are not equal. 



** Distribution of Patient Final Status over Different Treatment
```{r}
final_status = bladder %>%  
  select(id,treatment,recur,status) %>% 
  group_by(treatment,status) %>% summarise(n = n())

recurr_df2 = data.frame(final_status)

ggplot(data=recurr_df2, aes(x=as.factor(status), y= n,fill=as.factor(status))) + geom_bar(stat="identity", position=position_dodge())+geom_text(aes(label = n),color = "white",hjust=0.2,vjust=1.5,size=3,position = position_dodge(0.9))+facet_grid(. ~ treatment) + labs(title="Distribution of Patient Final Status over Different Treatment", x = "Status", y = "Counts", fill = "Status")
```


*Distribution of Recurrences Patients Final Status of over Treatment
```{r}
final_status_R = bladder %>% 
  filter(recur > 0) %>% 
  select(id,treatment,recur,status) %>% 
  group_by(treatment,status) %>% summarise(n = n())

final_status_R = data.frame(final_status_R)

ggplot(data=final_status_R, aes(x=as.factor(status), y= n,fill=as.factor(status))) + geom_bar(stat="identity", position=position_dodge())+geom_text(aes(label = n),color = "white",hjust=0.2,vjust=1.5,size=3,position = position_dodge(0.9))+facet_grid(. ~ treatment) + 
labs(title="Distrubition of Final Status of Recurrence Patients", x = "status", y = "counts", fill = "status")
```



```{r}
#distribution of number of tumors and the largest initial tumor
d1 =  bladder %>% select(id,treatment,number,size) %>% group_by(treatment)

ggplot(d1, aes(x = number, y = size, color = treatment)) + 
  geom_point(alpha = .5,position= "jitter") +
  xlim(1,8) +
  ylim(1,8) +
  facet_grid(. ~ treatment)

```

- Examine whether number of tumors/size of tumors has something to do with number of recurrences
```{r}
d4 = bladder %>% filter(recur != 0) %>% 
     group_by(id,treatment,recur,status,number,size) %>% summarise(n = n())
d4 = data.frame(d4)

ggplot(d4, aes(x = number, y = size, color = recur)) + 
  geom_point(alpha = .5, position = "jitter") + 
  facet_grid(. ~ treatment)
```


- Examine whether number of tumors/size of tumors has something to do with treatment assignment -- randomized, so probably not
```{r}
d5 =bladder %>% group_by(treatment,recur,status,number,size) %>% summarise(n = n())

#boxplot,y=number of tumors
ggplot(d5, aes(x = treatment, y = number)) + geom_boxplot()+ geom_jitter(shape=16, position=position_jitter(0.2))
```

tumor size/number with final status
```{r}
d6 = bladder %>% group_by(id,treatment,status,size,number) %>% summarise(n = n())

ggplot(d6, aes(x = number, y = size, color = factor(status))) + 
  geom_point(alpha = .5, position = "jitter") + 
  facet_grid(. ~ treatment)
```


```{r}
d7=bladder%>%select(treatment,recur,status,number,size)
pairs(d7)
```

