---
title: |
  | Evaluating and Extending Three-Arm
  | Study in the Treatment of Superficial
  | Bladder Cancer
  |
  | P8108 Final Project, Fall 2022
  | 
author: |
  | Yunyi Jiang, Xiao Ma, Waveley Qiu, 
  | Xuehan Yan, Congyang Xie
  | 
date: "2022-12-05"
output:
  beamer_presentation:
    colortheme: "dolphin"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(survival)
library(tidyverse)
library(ggfortify)
```

# Background

## Superficial Bladder Cancer

- Also known as Stage 1 bladder cancer
- Common diagnosis (75% of bladder cancer cases$^1$) and rarely life-threatening on it own.
- Thought to arise due to urinary issues$^1$ or through "abnormalities of tryptophan metabolism"$^2$
- Particular interest in preventing recurrence of disease
  - Natural history study conducted in Sweden saw that "death was directly related to tumor grade, number of tumors, and volume of recurrences."$^3$

## Pyridoxine and Thiotepa
- _Pyridoxine_ (vitamin $B_6$) thought to reduce "abnormalities of tryptophan metabolism"
- _Thiotepa_ has been the standard of care for the treatment of superficial bladder cancers.
- Effects of these two therapies compared in randomized clinical trial conducted by Byar and Blackard in 1977

## Byar and Blackard (1977)
- Primary clinical interest: prevent and reduce recurrence of Stage 1 bladder cancer.
- Event time agnostic analysis conducted included comparing overall rates and percentages of occurrence between groups
  - Pairwise difference detected in rate of recurrence between thiotepa and placebo, and thiotepa and pyridoxine 
  - No other differences in event incidence between groups detected
- Survival analysis conducted involved the construction of life-table estimates.
  - Life-table analysis indicate that the time to first recurrence was significantly different between pyridoxine and placebo groups.
  - Analysis restricted to subjects who experienced recurrence after at least 10 months of follow-up

# Proposed Project

## Motivation
- Research ought to be reproducible, especially if data are open-source
- Different models may be more informative than the actuarial curves constructed

## Analysis Plan
1. Reproduce analysis conducted in original study to see if results can be replicated
2. Construct models that differ from original study to see if they can be more informative
  - Non-parametric: KM
  - Semi-parametric: Cox Proportional Hazards Model
  - Parametric: Weibull and Exponential Models

## Exploratory Data Analysis

```{r, echo=FALSE, message=FALSE, warning=FALSE}
data(cancer, package = "survival")
head(bladder1) %>% knitr::kable()
```

- `id`:	Patient id
- `treatment`:	Placebo, pyridoxine (vitamin B6), or thiotepa
- `recur`:	Number of recurrences
- `start`, `stop`:	The start and end time of each time interval
- `status`:	End of interval code, 0=censored, 1=recurrence, 2=death from bladder disease, 3=death from other/unknown cause

# Reproduction Analysis

## 

```{r, echo=FALSE}
# table 1
table1 <- tibble(treatment = c("Placebo", "Pyridoxine", "Thiotepa"),
                 `Evaluable Patients` = c(48,32,38),
                 `Number without Recurrence` = c(19,17,20),
                 `Number with Recurrence` = c(29,15,18),
                 `Total Recurrences` = c(87,57,44),
                 `Total Months of Follow-up` = c(1528,993,1183)
                 ) %>% 
  mutate(`Percent Recurrence` = round(num_with_recur/evaluable_patients*100,2),
         `Recurrence Rate` = round(total_recur/total_month_follow*100,2))

table1 %>% pivot_longer(cols = -1) %>% pivot_wider(names_from = "treatment", values_from = "value") %>% knitr::kable()
```

## Exploratory Data Anaysis

```{r, echo=FALSE}
dat = bladder1 %>% filter(start == 0 & status %in% c(0,1))

bladder_fit <- survfit(Surv(stop, status == 1) ~ treatment, data = dat)
bladder_fit %>% autoplot() + xlab("Time") + ylab("Survival Function")
```


## References

1. Superficial bladder cancer. Division of Urologic Surgery. (n.d.). Retrieved from https://urology.wustl.edu/urologic-cancers/bladder-cancer/surgery-for-superficial-b/

2. C;, B. D. B. (n.d.). Comparisons of placebo, pyridoxine, and topical thiotepa in preventing recurrence of stage I bladder cancer. Urology. Retrieved from https://pubmed.ncbi.nlm.nih.gov/414402/ 

3. Pasin, E., Josephson, D. Y., Mitra, A. P., Cote, R. J., &amp; Stein, J. P. (2008). Superficial bladder cancer: An update on etiology, molecular development, classification, and natural history. Reviews in urology. Retrieved December 3, 2022, from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2312342/ 







